Secure Composition of Robust and Optimising Compilers

### 2023-07-10 popl/main.tex
sec 3

This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
Then, this section introduces monitors for each of the properties presented earlier (\Cref{subsec:monitors}).

@@

\gls{sms} disregards out of bounds accesses:

@@

Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
Note, however, that this definition does not entail memory-safety issues introduced by side-channels, such as speculation.

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
\end{definition}

@@

\gls{cct} is a hyperproperty and, thus, not as easy to check with just monitors. %Beutner et al
Hence, the following defines \gls{scct}, a stricter variant that enforces the policy that no secret should appear on a trace.

@@

\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
  $
  \msscctsafe:=\left\{\trace \left| \begin{array}{rcl}
                                      \trace\in\mssafe & \text{ and } & \trace\in\scctsafe
                                    \end{array}\right.\right\}
  $
\end{definition}

@@

Monitors can be used to enforce safety properties and this section presents concrete instantiations of monitors for the previousvy presented properties (\Cref{subsec:propdefs}).
To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation performs the actual work of projecting the events, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

The state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
This restriction effectively disallows reallocation to reassign the same location to an object.
However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
In the reflexive case of $\monitorcheck{\tmsmonitor}{\tmsmonitor}{\absevent}$, the monitor $\tmsmonitor$ must not contain allocated locations, i.e., $\tmsmonitor.allocated=\emptyset$.
This ensures that programs do not forget to free memory.

@@

  return y;
}
  \end{lstlisting}
  Relating the trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Dealloc}\ \loc[_{1}]\ \comp$.
  It is easy to see that $\montmssafe{\abstrace}$.
  From \Thmref{lem:mon:tmsafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}.
\end{example}

@@

\begin{example}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
  For the sake of this example, the pointers have been allocated by the component and passed to the context.
  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
  Because of this, there is an out of bounds memory access right before exiting the function.
  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.
  It is easy to see that, after the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
  Furthermore, all uses up to the last are acceptable by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}. Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls{sms}.
\end{example}

@@

\begin{example}[Data independent timing mode]
  Consider an excerpt from a component that reads from non-confidential memory \texttt{x} and writes the result to confidential memory \texttt{y}.
  The trace of just this part may look like $\ev{Use}\ \loc_{x}\ 1729;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
  In terms of the abstract monitor trace, this is just $\absev{Any}$.
  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace can contain $\ev{Use}\ \loc_{y}\ 1729;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 1729;\comp\lock$.
  Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
\end{example}

@@

The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.%
\MP{?canetti?} %% protocols under certain security notions are not secure when composed
The sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is defined as follows: given an $\src{L}$ program $\src{p}$, its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

\subsubsection{Securing Optimisations}
Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in practice~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MPin{survey to cite? this is a bit informal MK: hard to find a survey on different compiler passes... I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
Consider the compilation passes of LLVM~\cite{lattner2004llvm}.
In LLVM, compiler passes are handled in the \texttt{PassManager} class.
By means of manipulating the internal state of the \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
That way, an order of optimisations can be found by hand that yields the best results.
Take \gls{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls{dce}, which removes dead code by rewriting conditional branches.
Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
let a = true in
if a then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
if true then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
print "a"
\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
      % edges
      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls{dce}} (progunopt);
      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls{cf}} (progoptcf);
      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls{dce}} (progoptcfdce);
    \end{tikzpicture}
  \end{center}
  \caption{Example program where the level of optimisations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
\end{figure}
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.

@@

\subsubsection{Proof Sketch of Robust Safety Property Preservation}

The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature that are also used for similar theorems in this paper.

When proving robust safety property preservation for the concrete property \gls{tms}, a target-level execution of a compiled $\Ltms$ component $\src{\library_{\comp}}$ linked with an arbitrary context $\trg{\library_{\ctx}}$ is assumed, where the execution yields a trace $\trg{\trace}$.
The proof goal is to show that $\trg{\trace}$ satisfies \gls{tms}.
To get this satisfaction, one way of doing so is to use the other assumption in the proof context: $\rsat{\src{\library_{\comp}}}{\tmssafe}$.
However, this requires a source-level execution for some concrete $\src{\library_{\ctx}}$ that yields a trace $\src{\trace}$ similar to $\trg{\trace}$.
This paper uses the well-known trace-based backtranslation technique~\cite{abate2019jour} to construct such a context $\src{\library_{\ctx}}$.
It is shown for that obtained context $\src{\library_{\ctx}}$ that the whole program $\link{\src{\library_{\ctx}}}{\src{\library_{comp}}}$ is well-typed.
By \Cref{thm:wt:tms}, the fact that this program satisfies \gls{tms} is obtained and by typesafety, it is known that there must be some $\Ltms$ execution.
The proof now needs to relate the two execution with eachother.
\Cref{fig:proofdiag:rtp} visualizes this situation.
At the bottom, there is the target-level execution between operational states $\trg{\emptyset},\trg{\Omega_{1}}$, and so on.
The execution starts in the context $\trg{\library_{\ctx}}$ and is split by transitivity of $\trg{\estepto{}{}{}}$ into parts before, in, and after the component $\cc{\Ltms}{\trg{L}}(\src{p})$ executes\footnote{For sake of simplicity, this paper considers only one context-switch from context to component and back again. Recursive calls between context and component functions are disregarded.}.
Initially, the empty operational states $\src{\emptyset}$ and $\trg{\emptyset}$ need to agree by means of a cross-language state relation $\multimap_{\emptyset}$.
This relation is indexed by a function $\delta$ that maps $\trg{L}$-level locations $\trg{\loc}$ to $\Ltms$-level locations.
The operational state of the executions up to the component $\src{p}$ need to agree via $\multimap_{\delta}$, both contexts need to behave in a similar way such that the traces also agree with each other.
However, the relation allows $\trg{\library_{\ctx}}$ to perform internal actions that may change the operational state but are not visible on the trace, such as internal function calls, that does not $\src{\library_{\ctx}}$ need to reflect.
Before and after the compiler wrapper code, the operational states must agree via another cross-language relation $\approx_{\delta_{w_{}}}$, which disallows these mismatches with respect to internal function calls.
Intuitively, the behavior of the components should be tight: the source-level execution must be preserved by the compiler in the target-level execution.
Dually, when the component is done, the relations take turns again and the states must be related with $\multimap_{\delta_{w_{2}}}$
Finally, since both execution traces are similar in the sense that their projections $\msfilterLtms{\src{\trace}}$ and $\msfilterL{\trg{\trace}}$ yield specification traces that satisfy \Cref{def:monsat}, the claim follows \Cref{lem:mon:tmsafe}.

\MPin{
	why can we use a BT?
	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
	all these intuitions are missing.
	they are technical, so we only keep them here.

  MK: but, we don't actually rewrite it to RSC when doing the proof! The compilers are not robustly preserving with respect to an arbitrary safety property, but, e.g., just for \gls{tms}!
}

>>>>

This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
Then, this section introduces monitors for each of the properties presented earlier (\Cref{subsec:monitors}).
\MP{why?}

@@

\gls{sms} disregards\MP{prohibits?} out-of-bounds accesses:

@@

Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
\end{definition}
\MPin{
	cite catalin's paper and the mswasm paper for the def of ms
}

@@

\gls{cct} is a hyperproperty\MP{hypersafety?}\MP{what is a hp?} and, thus, not as easy to check with just monitors.\MP{chy? cit?} %Beutner et al
Hence, we define \gls{scct}, a stricter variant that enforces the policy that no secret should appear on a trace.\MP{cit? did we take some inspiration}

@@

\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
  $
  \msscctsafe:=\left\{\trace \left| \begin{array}{rcl}
                                      \trace\in\mssafe & \text{ and } & \trace\in\scctsafe
                                    \end{array}\right.\right\}
  $
\end{definition}
\MPin{
	comment on the composition at this levle?
}

@@

Monitors can be used to enforce safety properties and this section presents concrete instantiations of monitors for the previousvy presented properties (\Cref{subsec:propdefs}).
\MPin{
	why
}
To lessen the burden when proving\MP{proving what}, each monitor makes use\MP{uses} of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation projects the events\MP{whaddaya mean?}, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.\MP{unclear sentence}
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

For \gls{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
This restriction effectively disallows reallocation to reassign the same location to an object.
However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
In the reflexive case of $\monitorcheck{\tmsmonitor}{\tmsmonitor}{\absevent}$, the monitor $\tmsmonitor$ must not contain allocated locations, i.e., $\tmsmonitor.allocated=\emptyset$.\MP{wut?}
This ensures that programs do not forget to free memory.

@@

  return y;
}
  \end{lstlisting}
  Relating the trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Dealloc}\ \loc[_{1}]\ \comp$.
  It is easy to see that $\montmssafe{\abstrace}$.
  From \Thmref{lem:mon:tmsafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}.
\end{example}
\MPin{
	example is good, but it'd be nice to tie it to the running example?
}

@@

\begin{example}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
  Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
  For the sake of this example, the pointers have been allocated by the component and passed to the context.
  The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
  At that stage, the induction variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}.
  Because of this, there is an out of bounds\MP{hyphenate: out-of-bounds} memory access right before exiting the function.
  This is also visible on the trace, which can be sketched as $\ldots\cdot\ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 2;\comp;\unlock\cdot\ldots$, where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
  Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is $\absev{Alloc}\ \loc_{x}\ \comp\ 2\cdot\absev{Alloc}\ \loc_{y}\ \comp\ 2\cdot\absev{Use}\ \loc_{x}\ \comp\ 0\cdot\absev{Use}\ \loc_{y}\ \comp\ 0\cdot\absev{Use}\ \loc_{x}\ \comp\ 1\cdot\absev{Use}\ \loc_{y}\ \comp\ 1\cdot\absev{Use}\ \loc_{x}\ \comp\ 2$.

  It is easy to see that,\MP{erase. we don't know what's easy} after the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
  Furthermore, all uses up to the last are acceptable by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}. Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls{sms}.
\end{example}

@@

\begin{example}[Data-independent timing mode]
  Consider an excerpt from a component that reads from non-confidential memory \texttt{x} and writes the result to confidential memory \texttt{y}.
  The trace of just this part may look like $\ev{Use}\ \loc_{x}\ 1729;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 1729;\comp;\lock$.
  In terms of the abstract monitor trace, this is just $\absev{Any}$.
  Running this on the monitor would result in getting stuck, since there is no matching rule to step in the presence of $\absev{Any}$ event. % pun intended
  By means of additional features to ensure cryptographic constant time even in the presence of memory reads and loads, such as a flag to enable a data independent timing mode, which is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors, the original trace can contain $\ev{Use}\ \loc_{y}\ 1729;\comp;\unlock$ instead of $\ev{Use}\ \loc_{y}\ 1729;\comp\lock$.
  Because of this, the whole trace of the component equates to $\absev{\emptyevent}$ and the monitor can step without getting stuck.
\end{example}

@@

The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.%
\MP{?canetti?} %% protocols under certain security notions are not secure when composed
The sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is defined as follows: given an $\src{L}$ program $\src{p}$, its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
\MPin{
	section 3.1 never defined the $\cap$ operator on properties, nor on classes.
	it uses the 'and'.
	rectify
}

@@

\subsubsection{Securing Optimisations}
Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
% \MPin{survey to cite? this is a bit informal MK: hard to find a survey on different compiler passes... I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
Consider the compilation passes of LLVM~\cite{lattner2004llvm}.
In LLVM, compiler passes are handled in the \texttt{PassManager} class.
By means of manipulating the internal state of the \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
That way, an order of optimisations can be found by hand that yields the best results.
Take \gls{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls{dce}, which removes dead code by rewriting conditional branches.
Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
\begin{figure}[!ht]
  \begin{center}
    \begin{tikzpicture}
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
let a = true in
if a then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
if true then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
print "a"
\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
      % edges
      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls{dce}} (progunopt);
      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls{cf}} (progoptcf);
      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls{dce}} (progoptcfdce);
    \end{tikzpicture}
  \end{center}
  \caption{Example program where the level of optimisations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
\end{figure}
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.

@@

\subsubsection{Proof Sketch of Robust Temporal Memory Safety Preservation}

The proof of \Cref{thm:cca:rtp:tms} uses standard techniques in secure compilation literature and is visualized in \Cref{fig:proofdiag:rtp}.
\MPin{
	explain how to read the figure
}
Concretely, given the execution of the compiled $\Ltms$ program linked with an arbitrary $\trg{L}$ context yields a trace $\trg{\trace}$, it uses a trace-based backtranslation of $\trg{\trace}$ to reconstruct an $\src{L_{\tmssafe}}$ context.
\MPin{
	what does this mean?
	what is a backtranslation?
}
\MPin{
	why can we use a BT?
	it's because the def in section 2.2 is eq to anotehr one, because the props are Safety, we use the RSC equivalent def.
	all these intuitions are missing.
	they are technical, so we only keep them here.
}
\Cref{thm:wt:tms} already gives strong guarantees for programs written in $\src{L_{\tmssafe}}$ and can be exploited for the backtranslated context by ensuring that it typechecks.
Well-typing\MP{well-typedness} means the $\src{L_{\tmssafe}}$ program runs, so the rest of the proof is about relating the $\trg{L}$-level execution with the $\src{L_{\tmssafe}}$-level execution.

For the $\trg{L}$-execution, the compiler wrapper\MP{undefined terminology. call it like this before} code at state $\trg{\Omega_{1}}$ may either crash the program or continue execution $\trg{\Omega_{w_{1}}}$ right before control is transferred to the component.
In this specific case, the compiler needs not insert additional code after the component is done with its execution, but a symmetric situation can happen here.\MP{vague and imprecise}
Similarly, the backtranslation may also insert wrapper code prior and after switching to and from the component to the context. \MP{wut? why is a wrapper needed?}
\MKin{
  more detail or not? ;
  i.e., explain $\multimap$ and $\approx$?
  \MPin{
  yes
  }
}
\MPin{
	the general idea that later proofs use the same structure is also not here.
}
